### 2023-10-19 draft_jss/draft_jss.Rmd
clean tutorials parts

---
documentclass: jss
author:
  - name: FirstName LastName
    orcid: 0000-0000-0000-0000
    affiliation: University/Company
    # use this syntax to add text on several lines
    address: |
      | First line
      | Second line
    email: \email{name@company.com}
    url: https://posit.co
  - name: Second Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    address: |
      | Department of Statistics and Mathematics,
      | Faculty of Biosciences,
      | Universitat Autònoma de Barcelona
    affiliation: |
      | Universitat Autònoma 
      | de Barcelona
    # use a different affiliation in adress field (differently formated here)
    affiliation2: Universitat Autònoma de Barcelona
title:
  formatted: "A Capitalized Title: Something about a Package \\pkg{foo}"
  # If you use tex in the formatted title, also supply version without
  plain:     "A Capitalized Title: Something about a Package foo"
  # For running headers, if needed
  short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output:
  bookdown::pdf_book:
    base_format: rticles::jss_article
  bookdown::word_document2:
    default
knit: (function(inputFile, encoding){
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = here::here("output"), output_format = "all") })
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
```

@@

```{r rcpresentation, echo=FALSE, fig.cap="my caption", out.width = '90%', fig.cap="Reservoir computing is composed of an input layer, a reservoir and an output layer. Connection between input layer and reservoir and inside reservoir are random. Only the output layer is optimized based on a ridge penalized linear regression."}
knitr::include_graphics(here::here("reservoirnet : an R package for reservoir computing/images/schema.png/image1.png"))
```

@@

@@

Reservoir Computing (RC) is well suited to both regression and classification tasks. We will introduce a simple example of classification task. This chapter is largely inspired from reservoirnet vignette on
[classification](https://cran.r-project.org/web/packages/reservoirnet/vignettes/Classification_with_RC.html).

The first step is to load useful packages. For this example we will load both reservoirnet and ggplot2 (add ref) a useful package for datavisualisation.

```{r}
library(reservoirnet)
library(ggplot2)
```

@@

The Japanese vowel dataset is composed of 640 utterances of the Japanese vowel \ae, from 9 different male speakers. The goal of this task is to assign to each utterance the label of its speaker. Dataset is split between a 270 utterances training set and a 340 utterances testing set. Each spoken utterance is a timeseries of 7\~29 timesteps. Each timestep of signal is a 12 dimensional vector representing Linear Prediction Coefficient (LPC), which encode the audio signal into the cepstral domain (a variant of the frequency domain).

References M. Kudo, J. Toyama and M. Shimbo. (1999). "Multidimensional
Curve Classification Using Passing-Through Regions". Pattern Recognition
Letters, Vol. 20, No. 11--13, pages 1103--1111.
<https://archive.ics.uci.edu/ml/machine-learning-databases/JapaneseVowels-mld/>

The data set is splitted into a train and a test set.

```{r}
japanese_vowels <- reservoirnet::generate_data(dataset = "japanese_vowels")$japanese_vowels

```{r fig.cap="Vowel dataset"}
sample_per_speaker <- 30
n_speaker <- 9
X_train_per_speaker <- list()

for (i in 0:8) {
 X_speaker <- X_train[((i*sample_per_speaker)+1):((i+1)*sample_per_speaker)]
 X_train_per_speaker[[i+1]]<-(as.numeric(unlist(sapply(X_speaker, t))))
}

d <- data.frame(LPC = unlist(X_train_per_speaker), 
                Speaker = factor(rep(1:9,times = sapply(X_train_per_speaker,length))))

# plot data
ggplot(d,aes(x = Speaker, y = LPC)) + geom_boxplot()+theme_bw()
```

@@

As ReservoirPy Nodes are built to work on sequences, the simplest setup
to solve this task is sequence-to-sequence encoding, also called
transduction. A model is trained on encoding each vector of input
sequence into a new vector in the output space. Thus, a sequence of
audio yields a sequence of label, one label per timestep.

@@

#### Train a simple Echo State Network to solve this task:

@@

#[source >> reservoir, source] >> readout
model <- list(source %>>% reservoir, source) %>>% readout
```

#### Fit the model

@@

Get the scores:

There are 9 speakers, hence the output space is 9-dimensional. The
speaker label is the index of the output neuron with maximum activation.

```{r}
accuracy <- function(pred, truth){
  mean(pred == truth)
}
```

```{r}
Y_pred_class <- sapply(Y_pred, FUN = function(x) apply(as.matrix(x),1,which.max))
Y_test_class <- sapply(Y_test, FUN = function(x) apply(as.matrix(x),1,which.max))
score <- accuracy(array(unlist(Y_pred_class)), array(unlist(Y_test_class)))

@@

    ## [1] "Accuracy: 91.296%"

@@

We can create a more elaborated model where inference is performed only
once on the whole input sequence. Indeed, we only need to assign one
label to each input sequence. This new setup is known as a
sequence-to-vector model, and this is usually the type of model we refer
to when talking about classification of sequencial patterns.

@@

```{r}
source <- reservoirnet::createNode("Input")
readout <- reservoirnet::createNode("Ridge",ridge=1e-6)
reservoir <- reservoirnet::createNode("Reservoir", units = 500, lr=0.1, sr=0.9)

#source >> reservoir >> readout
model <- source %>>% reservoir %>>% readout
```

We need to modify the training loop by hand a bit to perform this task:

first, we compute all reservoir states over the input sequence using the
reservoir.run method. then, we gather in a list only the last vector of
the states sequence.

@@

We can now train the readout only on the last state vectors. Here,
Y_train is an array storing a single label for each utterance.

```{r}
res <- reservoirnet::reservoirR_fit(readout,X = states_train, Y = Y_train)
summary(res)
```

    ## Parametrs using to fit:
    ##  warmup: 0 ; stateful: FALSE ; reset: FALSE 
    ## results of fitting:
    ## 'Ridge-5': Ridge(ridge=1e-06, input_bias=True, in=500, out=9)

We also modify the inference code using the same method as above:

@@

@@

    ## [1] "Accuracy: 88.108%"

@@

Vignette:
<https://cran.r-project.org/web/packages/reservoirnet/vignettes/basic_usage_01.html>

# Study case: Prediction covid-19

@@

4.1) Study case

@@

4.2) Evaluation framework

-   Train and test set

@@

4.3) Hyperparameter optimisation using GA

@@

4.4) Present results from GA

@@

5)  Discussion and conclusion

<!-- ## Code formatting -->

<!-- In general, don't use Markdown, but use the more precise LaTeX commands instead: -->

<!-- * \proglang{Java} -->

<!-- * \pkg{plyr} -->

<!-- One exception is inline code, which can be written inside a pair of backticks (i.e., using the Markdown syntax). -->

<!-- If you want to use LaTeX commands in headers, you need to provide a `short-title` attribute. You can also provide a custom identifier if necessary. See the header of Section \ref{r-code} for example. -->

<!-- # \proglang{R} code {short-title="R code" #r-code} -->

<!-- Can be inserted in regular R markdown blocks. -->

<!-- ```{r} -->

<!-- x <- 1:10 -->

<!-- x -->

<!-- ``` -->

<!-- ## Features specific to \pkg{rticles} {short-title="Features specific to rticles"} -->

<!-- * Adding short titles to section headers is a feature specific to \pkg{rticles} (implemented via a Pandoc Lua filter). This feature is currently not supported by Pandoc and we will update this template if [it is officially supported in the future](https://github.com/jgm/pandoc/issues/4409). -->

<!-- * Using the `\AND` syntax in the `author` field to add authors on a new line. This is a specific to the `rticles::jss_article` format. -->

>>>>

---
documentclass: jss
bibliography: article.bib
author:
  - name: FirstName LastName
    orcid: 0000-0000-0000-0000
    affiliation: University/Company
    # use this syntax to add text on several lines
    address: |
      | First line
      | Second line
    email: \email{name@company.com}
    url: https://posit.co
  - name: Second Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation'
    # To add another line, use \AND at the end of the previous one as above
  - name: Fourth Author
    orcid: 0000-0000-0000-0000
    address: |
      | Department of Statistics and Mathematics,
      | Faculty of Biosciences,
      | Universitat Autònoma de Barcelona
    affiliation: |
      | Universitat Autònoma 
      | de Barcelona
    # use a different affiliation in adress field (differently formated here)
    affiliation2: Universitat Autònoma de Barcelona
title:
  formatted: "Reservoir computing in R : a tutorial using \\pkg{reservoirnet}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Reservoir computing in R : a tutorial using reservoirnet"
  # For running headers, if needed
  short:     "\\pkg{reservoirnet}: reservoir computing in R"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output:
  bookdown::pdf_book:
    base_format: rticles::jss_article
  bookdown::word_document2:
    default
knit: (function(inputFile, encoding){
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = here::here("draft_jss"), output_format = "all") })
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
set.seed(1)
```

@@

```{r rcpresentation, echo=FALSE, fig.cap="my caption", out.width = '90%', fig.cap="Reservoir computing is composed of an input layer, a reservoir and an output layer. Connection between input layer and reservoir and inside reservoir are random. Only the output layer is optimized based on a ridge penalized linear regression."}
knitr::include_graphics("images/schema.png/image1.png")
```

@@

Reservoir Computing (RC) is well suited to both regression and
classification tasks. We will introduce a simple example for both task.

@@

This example is largely inspired from the [classification tutorial of reservoirpy](https://github.com/reservoirpy/reservoirpy/blob/master/tutorials/5-Classification-with-RC.ipynb). The first step is to load useful packages. For this example we will load both reservoirnet, ggplot2 [@wickham_ggplot2_2016] and dplyr
[@wickham_dplyr_2023].

```{r message=FALSE, warning=FALSE}
library(reservoirnet)
library(ggplot2)
library(dplyr)
```

@@

To illustrate the classification task, we will use the Japanese vowel dataset (@kudo_multidimensional_1999). The data can be loaded from reservoirnet as follow :

```{r}
japanese_vowels <- reservoirnet::generate_data(dataset = "japanese_vowels")[[1]]
X_train <- japanese_vowels$X_train
Y_train <- japanese_vowels$Y_train
X_test <- japanese_vowels$X_test
Y_test <- japanese_vowels$Y_test
```

The dataset is composed of 640 utterances of the Japanese vowel \ae, from 9 different speakers. The goal is to assign each uterance to the its speaker. Each utterances is a time series of 7 to 29 timesteps encoded by a 12 dimensional vector representing Linear Prediction Coefficient (LPC).

```{r fig.cap="Vowel dataset, sample with 3 speakers and 2 utterance each."}
vec_sample <- c(1, 2, 41, 42, 71, 72)
dfplot_vowel <- lapply(vec_sample,
                 FUN = function(i){
                   speaker <- which(Y_test[[i]] == 1)
                   X_test[[i]] %>%
                     as.data.frame() %>%
                     tibble::rowid_to_column(var = "Time") %>%
                     tidyr::pivot_longer(cols = -Time,
                                         names_to = "component",
                                         values_to = "LPC") %>%
                     mutate(speaker = speaker, .before = 1,
                            uterrance = i) %>%
                     return()
                 }) %>%
  bind_rows()

ggplot(dfplot_vowel, mapping = aes(x = Time, y = LPC, color = component)) +
  geom_line() +
  facet_wrap(uterrance ~ speaker,
             labeller = label_bquote(cols = "speaker" : .(speaker)),
             ncol = 2) +
  theme_minimal() +
  theme(legend.position = "none")

```

@@

We can first solve this task using a sequence-to-sequence encoding approach. For this task, the goal is to predict the speaker for each time step of each utterance.

The first step is to get the data where the label is repeated for each time step. This is easily done with the `repeat_targets` argument as follow :

@@

Then we can train a simple Echo State Network to solve this task. Here we will connect both the input layer and the reservoir layer to the readout layer.

@@

model <- list(source %>>% reservoir, source) %>>% readout
```

We can then fit the model and predict the labels for the test data :

@@

From the `Y_pred` and `Y_test` we can represent graphically the forecast for some patients :

```{r fig.cap="Prediction in a sequence-to-sequence approach 6 samples with 3 speakers and 2 utterance each. The higher the score of the speaker, the lighter the color.", fig.height=6}
dfplotseqtoseq <- lapply(vec_sample,
                 FUN = function(i){
                   speaker <- which(Y_test[[i]][1,] == 1)
                   Y_pred[[i]] %>%
                     as.data.frame() %>%
                     tibble::rowid_to_column(var = "Time") %>%
                     tidyr::pivot_longer(cols = -Time,
                                         names_to = "pred_speaker",
                                         values_to = "prediction") %>%
                     mutate(pred_speaker = gsub(x = pred_speaker,
                                                pattern = "V", "")) %>%
                     mutate(speaker = speaker, .before = 1,
                            uterrance = i) %>%
                     return()
                 }) %>%
  bind_rows()

ggplot(dfplotseqtoseq, mapping = aes(x = Time,
                                     y = pred_speaker,
                                     fill = prediction)) +
  geom_tile() +
  facet_wrap(uterrance ~ speaker,
             labeller = label_bquote(cols = "speaker" : .(speaker)),
             ncol = 2) +
  theme_minimal() +
  labs(y = 'Predicted speaker',
       fill = "Prediction score")
```

For those 6 utterances, the model correctly identify the speaker for most of the time steps. We can then evaluate the overall accuracy of the model :

```{r}
accuracy <- function(pred, truth) mean(pred == truth)

Y_pred_class <- sapply(Y_pred, FUN = function(x) apply(as.matrix(x),
                                                       1,
                                                       which.max))
Y_test_class <- sapply(Y_test, FUN = function(x) apply(as.matrix(x),
                                                       1,
                                                       which.max))
score <- accuracy(array(unlist(Y_pred_class)), array(unlist(Y_test_class)))

@@

@@

Another approach is the sequence-to-vector encoding. For this task we aim to predict the speaker of the whole utterance instead of the speaker of each time step (i.e the label is assign to the whole sequence).

@@

In this example, we only connect the reservoir to the output layer.

```{r}
reservoir <- reservoirnet::createNode("Reservoir", units = 500, lr=0.1, sr=0.9)
readout <- reservoirnet::createNode("Ridge",ridge=1e-6)
```

To perform this task, we need to modify the training and testing process. Thanks to the echo property of RC, the information of previous time step is stored in the reservoir (i.e RC have a "memory"). Therefore, the last vector of states accumulates information from all previous states. For this task of sequence-to-vector encoding we only use this last state. This performed as follow :

@@

Then we can train the readout based on this last state vector. In that case, `Y_train`

We can now train the readout only on the last state vectors. Here `Y_train` contains a single label for each utterance.

```{r}
res <- reservoirnet::reservoirR_fit(readout,X = states_train, Y = Y_train)
```

The prediction is also modified to use only the final state :

@@

```{r fig.cap="Prediction in a sequence-to-sequence approach 6 samples with 3 speakers and 2 utterance each. The higher the score of the speaker, the lighter the color."}
dfplotseqtovec <- lapply(vec_sample,
                 FUN = function(i){
                   speaker <- which(Y_test[[i]][1,] == 1)
                   Y_pred[[i]] %>%
                     as.data.frame() %>%
                     tidyr::pivot_longer(cols = everything(),
                                         names_to = "pred_speaker",
                                         values_to = "prediction") %>%
                     mutate(pred_speaker = gsub(x = pred_speaker,
                                                pattern = "V", "")) %>%
                     mutate(speaker = speaker, .before = 1,
                            uterrance = i,
                            target = speaker == pred_speaker) %>%
                     return()
                 }) %>%
  bind_rows()

ggplot(dfplotseqtovec,
       mapping = aes(x = pred_speaker,
                     y = prediction,
                     color = target)) +
  geom_point() +
  facet_wrap(uterrance ~ speaker,
             labeller = label_bquote(cols = "speaker" : .(speaker)),
             ncol = 2) +
  scale_color_manual(values = c("darkgrey", "red")) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y = 'Score',
       fill = "Speaker")
```

We observe, for those 6 uterrances that the model is correct for each of them. We can also compute the overall accuracy :

@@

@@

### Goal

Here you will learn basic use of reservoirnet package. This will be
illustrated with the covid-19 dataset included in the pakage. Those data
contain the hospitalisations, positive RT-PCR and overall RT-PCR from
Santé Publique France available on data.gouv.fr (see `help(dfCovid)` for
more informations). Our goal will be to forecast the number of
hospitalised patients 14 days ahead. To do so, we will first learn on
data prior to 2022-01-01 and forecast on the following data.

We can load the data and set the given task :

```{r}
data("dfCovid")
dist_forecast = 14
traintest_date = as.Date("2022-01-01")
```

### Covid-19 data

Because both RT-PCR have strong variations, we will first proceed by
taking the moving average on the last 7 days of those features. Also we
add an outcome and an `outcomeDate` colum which will be useful to train
the model. Furthermore we will compute the
`outcome_deriv = outcome - hosp` which corresponds to the variation of
hospitalisation compared to the current number of hospitalized patients.

```{r}
dfOutcome <- dfCovid %>%
  # outcome at 14 days
  mutate(outcome = lead(x = hosp, n = dist_forecast),
         outcomeDate = date + dist_forecast,
         outcome_deriv = outcome - hosp) %>%
  # rolling average for tested and positive_pcr
  mutate_at(.vars = c("Positive", "Tested"),
            .funs = function(x) slider::slide_dbl(.x = x,
                                                  .before = 6,
                                                  .f = mean))
```

We can now plot the data :

```{r fig.cap="Hospitalisations, IPTCC and positive PCR of Bordeaux University Hospital."}
dfOutcome %>%
  tidyr::pivot_longer(cols = c("hosp", "Positive", "Tested")) %>%
  ggplot2::ggplot(mapping = aes(x = date, y = value)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y") +
  theme_bw() +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  labs(color = "") +
  theme(legend.position = "bottom")
```

### First reservoir

Setting a reservoir is easily done with the `createNode()` function. The
important hyperparameters are the following :

-   Number of nodes (`units`) : it corresponds to the number of nodes
    inside the reservoir. Usually, the more the better but more nodes
    increase the computation time.
-   Leaking rate (`lr`) : the leaking rate corresponds to the balance
    between the new inputs and the previous state. A leaking rate of 1
    only consider informations from the new inputs.
-   Spectral radius (`sr`): the spectral radius is the maximum absolute
    eigenvalue of the reservoir connectivity matrix. A small spectral
    radius induces stable dynamics inside the reservoir, a high spectral
    radius induces chaotic regimen inside the reservoir.
-   Input scaling (`input_scaling`): the input scaling is a gain applied
    to the input features of the reservoir.

For this part of the tutorial, we will set the hyperparameter at a given
value. Hyperparameter opitmisation will be detailed at section [Study
case](#studycase).

```{r}
reservoir <- reservoirnet::createNode(nodeType = "Reservoir",
                                      seed = 1,
                                      units = 500,
                                      lr = 0.7,
                                      sr = 1,
                                      input_scaling = 1)
```

Then we can feed the data to the reservoir and see the activation
depending on time of the reservoir. To do so, we first prepare the data
and transform it to an array.

```{r}
## select explanatory and transform it to an array
X <- dfOutcome %>%
  filter(outcomeDate < traintest_date) %>%
  select(hosp, Positive, Tested) %>%
  as.matrix() %>%
  as.array()
```

Then we run the `predict_seq` function. The parameter reset is set to
TRUE to be sure that the node is clear from data (it is optional here).

```{r}
reservoir_state <- predict_seq(node = reservoir, X = X, reset = TRUE)
```

Now we can visualise node activation using the `plot` function :

```{r fig.cap="Node acivation over time."}
plot(reservoir_state)
```

We see a lot of nodes with a steady state. It will be really difficult
for the ridge output layer to learn on those nodes because they present
no information. The problem is caused by the different scales of the
features, to take this into account, we can divide them by their maximum
value to scale each of them between `-1` and `1` :

```{r}
stand_max <- function(x) return(x/max(x))
# scaled features
Xstand <- dfOutcome %>%
  filter(date < traintest_date) %>%
  select(hosp, Positive, Tested) %>%
  mutate_all(.funs = stand_max) %>%
  as.matrix() %>%
  as.array()
```

We then feed them to the reservoir and plot the node activation :

```{r fig.cap="Node acivation over time. Scaled features"}
# feed them to the reservoir
reservoir_state_stand <- predict_seq(node = reservoir,
                                     X = Xstand,
                                     reset = TRUE)
# plot the output
plot(reservoir_state_stand)
```

The reservoir dynamics now look much better as no node seem to get
saturated.

### Forecast

In order to train the reservoir, we should train the last layer which
linearly combines the neuron's output.

#### Set the ESN

First we define this last layer with a ridge penalty of `1e3`. This
hyperparameter can also be optimised which will be discussed at section
[Study case](#studycase).

```{r}
readout <- reservoirnet::createNode(nodeType = "Ridge",
                                    ridge = 1e3)
```

And we connect it to the previously defined reservoir :

```{r}
model <- reservoirnet::link(reservoir, readout)
```

The model is now ready we must now train the last layer.

#### Set the data

First we separate the train set on which we will learn the ridge
coefficients and the test set on which we will make the forecast. We
define the train set to be all the data before 2022-01-01 and the test
data to be all the data to have forecast both on train and test set.

```{r}
# train set
yTrain <- dfOutcome %>%
  filter(outcomeDate <= traintest_date) %>%
  select(outcome)
yTrain_variation <- dfOutcome %>%
  filter(outcomeDate <= traintest_date) %>%
  select(outcome_deriv)
xTrain <- dfOutcome %>%
  filter(outcomeDate <= traintest_date) %>%
  select(hosp, Positive, Tested)
# test set
xTest <- dfOutcome %>%
  select(hosp, Positive, Tested)
```

We now standardise with the same formula as seen before. We learn the
standardisation on the training set :

```{r}
# standardise based on training set values
ls_fct_stand <- apply(xTrain,
                      MARGIN = 2,
                      FUN = function(x) function(feature) return(feature/(max(x))))
```

And we apply it to both the train and test sets :

```{r results=FALSE}
xTrainstand <- xTrain
xTeststand <- xTest
lapply(X = names(ls_fct_stand),
       FUN = function(x){
         xTrainstand[,x] <<- ls_fct_stand[[x]](feature = xTrain[,x])
         xTeststand[,x] <<- ls_fct_stand[[x]](feature = xTest[,x])
         return()
       })
```

Finally we convert all those data to array :

```{r}
# convert to array
lsdf <- lapply(list(yTrain = yTrain,
                    yTrain_variation = yTrain_variation,
                    xTrain = xTrainstand,
                    xTest = xTeststand),
               function(x) as.array(as.matrix(x)))
```

#### Train the model and predict

Now, the easy part. We are going to train the reservoir with the train
set. To do so, we set a warmup of 30 days during which the data are
propagating into the reservoir but not used to fit the ridge layer.

```{r}
### train the reservoir ridge output
fit <- reservoirnet::reservoirR_fit(node = model,
                                    X = lsdf$xTrain,
                                    Y = lsdf$yTrain,
                                    warmup = 30,
                                    reset = TRUE)
```

Now that the ridge layer is trained, we can perform predictions as seen
before. We set the parameter `reset` to TRUE in order to clean the
reservoir from the data used by the training set.

```{r}
### predict with the reservoir
vec_pred <- reservoirnet::predict_seq(node = fit$fit,
                                      X = lsdf$xTest,
                                      reset = TRUE)
```

```{r fig.cap="Forecast"}
dfOutcome %>%
  mutate(pred = vec_pred) %>%
  na.omit() %>%
  ggplot(mapping = aes(x = outcomeDate)) +
  geom_line(mapping = aes(y = outcome,
                          color = "observed")) +
  geom_line(mapping = aes(y = pred,
                          color = "forecast")) +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  scale_color_manual(values = c("#3772ff", "#080708", "#df2935")) +
  theme_bw() +
  labs(color = "", x = "Date", y = "Hospitalisations")
```

We observe that the model forecast is not fully accurate, both on the
test set and the train set. In that case, one option could be to reduce
ridge penalisation to fit more closely the data, the optimisation of
ridge hyperparameter will be discussed at section [Study
case](#studycase). Another possibility is to ease the learning of the
algorithm by forecasting the variation of the hospitalisation instead of
the number of hospitalised patients. For that step, we will learn on the
`outcome_deriv` contained in `yTrain_variation` data which is defined
outcome as `outcome_deriv = outcome - hosp`.

```{r fig.cap="Covid-19 hospitalisations forecast. The model is either trained to forecast the number of hospitalisations (denoted Raw) or the variation of the hospitalisations compared to current level of hospitalisation (denoted Variation)"}
## Fit reservoir on outcome variation instead of raw outcome
fit2 <- reservoirnet::reservoirR_fit(node = model,
                                     X = lsdf$xTrain,
                                     Y = lsdf$yTrain_variation,
                                     warmup = 30,
                                     reset = TRUE)
## Get the forecast on the test set
vec_pred2_variation <- reservoirnet::predict_seq(node = fit2$fit,
                                                 X = lsdf$xTest,
                                                 reset = TRUE)
## Transform the outome variation forecast into hospitalisation forecast
vec_pred2 <- vec_pred2_variation + xTest$hosp
## Plot the results
dfOutcome %>%
  mutate(Raw = vec_pred,
         Variation = vec_pred2) %>%
  tidyr::pivot_longer(cols = c(Raw, Variation),
                      names_to = "Outcome_type",
                      values_to = "Forecast") %>%
  na.omit() %>%
  ggplot(mapping = aes(x = outcomeDate)) +
  geom_line(mapping = aes(y = outcome,
                          color = "observed")) +
  geom_line(mapping = aes(y = Forecast,
                          color = "Forecast")) +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  facet_wrap(Outcome_type ~ .,
             labeller = label_bquote(cols = "Outcome" : .(Outcome_type))) +
  scale_color_manual(values = c("#3772ff", "#080708", "#df2935")) +
  theme_minimal() +
  labs(color = "", x = "Date", y = "Hospitalisations")
```

We observe an improvement of the forecast compared to the forecast of
hospitalisation previously discussed. From there, many improvement can
be implemented including : frequent update of the model, leverage
additional features and hyperparameter optimisation. This notions will
be discussed in the next [section](#studycase).

# Study case: Prediction covid-19 {#studycase}

@@

## Study case

@@

## Evaluation framework

### Train and test set

@@

## Hyperparameter optimisation using GA

@@

## Present results from GA

@@

## Present forecast performance

## Discussion on use case

# Previous work

Those work were done previously, the setting of train/test and
hyperparameter optimisation were different and probably they should be
done again in a similar setting for relevant comparison.

## Raw outcome, first derivative or second derivative ?

See
<https://github.com/thomasferte/PredictCovid/blob/esn_update_report/reporting/02_esn/01_explore_esn_prediction/00_reporting/explore_esn.md>

## Raw feature or scaled feature ?

See
<https://github.com/thomasferte/PredictCovid/blob/esn_update_report/reporting/02_esn/01_explore_esn_prediction/00_reporting/explore_esn.md>

## Smoothed or not smoothed

See
<https://github.com/thomasferte/PredictCovid/blob/esn_update_report/reporting/02_esn/01_explore_esn_prediction/00_reporting/explore_esn.md>

## Number of nodes

See
<https://github.com/thomasferte/PredictCovid/blob/esn_update_report/reporting/02_esn/01_explore_esn_prediction/00_reporting/explore_esn_r_pipeline.md>

## Use of second derivative among the interesting features

See
<https://github.com/thomasferte/PredictCovid/blob/esn_update_report/reporting/02_esn/01_explore_esn_prediction/00_reporting/tune_esn2.md>

## Results of random sampling with R

See
<https://github.com/thomasferte/PredictCovid/blob/esn_update_report/reporting/02_esn/01_explore_esn_prediction/meeting_1801/master_report.md>

# Discussion and conclusion

# Bibliography
